title: "GTCI Tables"
author: "Emerson Johnston"
---

# Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE
)

library(tidyverse)
library(scales)
library(showtext)
library(ggtext)
library(ggrepel)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(gt)
library(gtExtras)
library(webshot2)
library(magick)
library(pdftools)

source("R/theme.R")
source("R/helpers.R")
source("R/tables.R")
source("R/chart_layers.R")
source("R/maps.R")
```

# Chapter 1 

### Table 1.01 - How Much of U.S. STEM Training Goes to International Students?
```{r table 006}
meta_d  <- get_meta("TAB101")

TAB101 <- build_change_table_wide("TAB101", meta_d)

save_table(TAB101, "tab_1_01", profile = "ultrawide", stub_width_px = 420)
```


# Chapter 2 

### Table 2.01 - Where International Students in the United States Come From
```{r table 007}
meta_007 <- get_meta("TAB201")

TAB201 <- build_region_table(
  name = "TAB201",
  meta = meta_007
)

save_table(tab_008, "tab_2_02", profile = "ultrawide", stub_width_px = 260)
```

### Table 2.02 - What Share of Each Country's International Students Are Enrolled in Key STEM Fields?
```{r table 008}
meta <- get_meta("TAB202")

field_files <- c(
  "Engineering"              = "TAB202a",
  "Math & Computer Science"  = "TAB202b",
  "Physical & Life Sciences" = "TAB202c"
)
field_labels <- names(field_files)

read_field_long <- function(name) {
  read_fig(name) %>%
    mutate(across(-Year, ~ suppressWarnings(as.numeric(na_if(as.character(.x), "-"))))) %>%
    pivot_longer(-Year, names_to = "Country", values_to = "Share") %>%
    filter(!is.na(Share))
}

long_all <- imap(field_files, ~ read_field_long(.x) %>% mutate(Field = .y)) %>%
  bind_rows()

summ <- long_all %>%
  filter(Year %in% c(2009, 2024)) %>%
  mutate(point = if_else(Year == 2009, "First", "Last")) %>%
  select(Field, Country, point, Share) %>%
  pivot_wider(names_from = point, values_from = Share) %>%
  filter(!is.na(First), !is.na(Last)) %>%
  mutate(First = First * 100, Last = Last * 100,
         Change = Last - First)          # percentage-point change

# keep only countries with both 2009 and 2024 values in all three fields
keep_countries <- summ %>% count(Country) %>% filter(n == 3) %>% pull(Country)
summ <- summ %>% filter(Country %in% keep_countries)

ord_cols <- as.vector(t(outer(field_labels, c("First", "Last", "Change"),
                              paste, sep = "_")))

country_order <- c(
  "China",
  "India",
  "South Korea", 
  "Canada", 
  "Japan", 
  "Taiwan", 
  "Saudi Arabia",
  "Mexico",
  "Vietnam", 
  "Brazil",
  "Turkey/Türkiye", 
  "Nigeria",
  "United Kingdom", 
  "Germany", 
  "Hong Kong",
  "Nepal",
  "Indonesia",
  "France", 
  "Colombia", 
  "Pakistan"
)

flag_slug <- c(
  "China"          = "china",
  "India"          = "india",
  "South Korea"    = "south-korea",
  "Canada"         = "canada",
  "Japan"          = "japan",
  "Taiwan"         = "taiwan",
  "Saudi Arabia"   = "saudi-arabia",
  "Mexico"         = "mexico",
  "Vietnam"        = "vietnam",
  "Brazil"         = "brazil-",       # file is literally brazil-.png
  "Turkey/Türkiye" = "turkey",
  "Nigeria"        = "nigeria",
  "United Kingdom" = "united-kingdom",
  "Germany"        = "germany",
  "Hong Kong"      = "hong-kong",
  "Nepal"          = "nepal",
  "Indonesia"      = "indonesia",
  "France"         = "france",
  "Colombia"       = "colombia",
  "Pakistan"       = "pakistan"
)

df_008 <- summ %>%
  pivot_wider(
    names_from  = Field,
    values_from = c(First, Last, Change),
    names_glue  = "{Field}_{.value}"
  ) %>%
  mutate(flag = paste0("flags/", recode(Country, !!!flag_slug), ".png")) %>%
  select(Country, flag, all_of(ord_cols)) %>%
  arrange(match(Country, country_order))

tab_008 <- build_field_share_table(df_008, "tab_008", meta, field_labels,
                                    change_type = "ppt")
save_table(tab_008, "tab_2_02")
```

### Table 2.03 - Who Enrolls in U.S. Graduate Programs
```{r tab_009}
meta <- get_meta("TAB203")

cite_files <- c(
  "Science"     = "TAB203a",
  "Engineering" = "TAB203b",
  "Health"      = "TAB203c"
)

read_cite_long <- function(name) {
  read_fig(name) %>%
    mutate(across(-Year, ~ suppressWarnings(as.numeric(na_if(as.character(.x), "-"))))) %>%
    pivot_longer(-Year, names_to = "Group", values_to = "Count") %>%
    filter(!is.na(Count))
}

long_all <- imap(cite_files, ~ read_cite_long(.x) %>% mutate(Field = .y)) %>%
  bind_rows() %>%
  mutate(Group = recode(Group,
    "U.S. citizens and permanent residents" = "USC",
    "Temporary visa holders"                = "TVH"))

first_year <- min(long_all$Year, na.rm = TRUE)
last_year  <- max(long_all$Year, na.rm = TRUE)

# counts -> within-field citizenship share, per Field × Year
shares <- long_all %>%
  filter(Year %in% c(first_year, last_year)) %>%
  group_by(Field, Year) %>%
  mutate(Share = Count / sum(Count) * 100) %>%
  ungroup()

summ <- shares %>%
  mutate(point = if_else(Year == first_year, "First", "Last")) %>%
  select(Field, Group, point, Share) %>%
  pivot_wider(names_from = point, values_from = Share) %>%
  mutate(Change = (Last - First) / First * 100)   # relative change in share

ord_cols <- as.vector(t(outer(c("USC", "TVH"),
                              c("First", "Last", "Change"), paste, sep = "_")))

field_order <- c("Science", "Engineering", "Health")

df_009 <- summ %>%
  pivot_wider(
    names_from  = Group,
    values_from = c(First, Last, Change),
    names_glue  = "{Group}_{.value}"
  ) %>%
  select(Field, all_of(ord_cols)) %>%
  arrange(match(Field, field_order))

tab_009 <- build_citizenship_table(df_009, "tab_009", meta,
                        first_year = first_year, last_year = last_year,
                        value_is_share = TRUE,
                        footnote = paste(
                          "Within-field citizenship share of U.S. graduate enrollment in",
                          "science, engineering, and health. Change is the relative change",
                          "in share between", first_year, "and", last_year, "."))

save_table(tab_009, "tab_2_03")
```

### Table 2.04 - Who Earns U.S. STEM Doctorates
```{r table_010}
meta <- get_meta("TAB204")

phd_files <- c(
  "USC" = "TAB204a",
  "TVH" = "TAB204b"
)

read_phd_long <- function(name) {
  read_fig(name) %>%
    mutate(across(-Year, ~ suppressWarnings(as.numeric(na_if(as.character(.x), "-"))))) %>%
    pivot_longer(-Year, names_to = "Field", values_to = "Count") %>%
    filter(!is.na(Count)) %>%
    mutate(Field = recode(Field, "Social Science" = "Social Sciences"))
}

long_all <- imap(phd_files, ~ read_phd_long(.x) %>% mutate(Group = .y)) %>%
  bind_rows()

first_year <- min(long_all$Year, na.rm = TRUE)
last_year  <- max(long_all$Year, na.rm = TRUE)

summ <- long_all %>%
  filter(Year %in% c(first_year, last_year)) %>%
  mutate(point = if_else(Year == first_year, "First", "Last")) %>%
  select(Field, Group, point, Count) %>%
  pivot_wider(names_from = point, values_from = Count) %>%
  mutate(Change = (Last - First) / First * 100)

# wide counts: USC_First/Last/Change, TVH_First/Last/Change
counts_wide <- summ %>%
  pivot_wider(
    names_from  = Group,
    values_from = c(First, Last, Change),
    names_glue  = "{Group}_{.value}"
  )

# international share = TVH / (TVH + USC), per year, + relative change
df_010 <- counts_wide %>%
  mutate(
    SHR_First  = TVH_First / (TVH_First + USC_First) * 100,
    SHR_Last   = TVH_Last  / (TVH_Last  + USC_Last)  * 100,
    SHR_Change = (SHR_Last - SHR_First) / SHR_First * 100
  ) %>%
  select(Field,
         USC_First, USC_Last, USC_Change,
         TVH_First, TVH_Last, TVH_Change,
         SHR_First, SHR_Last, SHR_Change) %>%
  arrange(Field == "Non-sciences", desc(TVH_Last))

tab_010 <- build_citizenship_table(df_010, "tab_010", meta,
                        first_year = first_year, last_year = last_year,
                        value_is_share = FALSE,
                        share_group = TRUE,
                        footnote = paste(
                          "Research doctorate recipients by field and citizenship.",
                          "Change is the percent change between", first_year, "and", last_year, ".",
                          "International share is temporary visa holders as a percentage of",
                          "all recipients in the field; its change is relative.",
                          "Fields ordered by 2024 international count; non-sciences shown for comparison."))

save_table(tab_010, "tab_2_04", profile = "ultrawide", stub_width_px = 240)
```


# Chapter 3 

### Table 3.01 - Which STEM Graduates Work in STEM Occupations
```{r tab_14}
meta <- get_meta("TAB301")   # or reuse TAB301's meta

df_029t <- read_fig("TAB301") %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Level", values_to = "Share") %>%
  filter(!is.na(Share), Year %in% c(1993, 2003, 2013, 2023)) %>%
  mutate(
    Share = Share * 100,
    Level = case_when(
      str_detect(Level, regex("bach", ignore_case = TRUE)) ~ "Bachelors",
      str_detect(Level, regex("mast", ignore_case = TRUE)) ~ "Masters",
      str_detect(Level, regex("doct", ignore_case = TRUE)) ~ "Doctorate"
    )
  ) %>%
  pivot_wider(names_from = Year, values_from = Share, names_prefix = "Y") %>%
  mutate(
    Chg2003 = (Y2003 - Y1993) / Y1993 * 100,
    Chg2013 = (Y2013 - Y2003) / Y2003 * 100,
    Chg2023 = (Y2023 - Y2013) / Y2013 * 100,
    Change  = (Y2023 - Y1993) / Y1993 * 100,
    Occupation = factor(Level, levels = c("Bachelors", "Masters", "Doctorate"))
  ) %>%
  arrange(Occupation) %>%
  select(Occupation, Y1993, Chg2003, Chg2013, Chg2023, Y2023, Change)

tab_014 <- build_occupation_share_table(
  df_029t, "tab_014", meta,
  level_cols    = c("Y1993", "Y2023"),
  chg_cols      = c("Chg2003", "Chg2013", "Chg2023"),
  spanner_label = "Share employed in STEM",
  footnote = paste(
    "1993 is the share of graduates at each degree level employed in a STEM",
    "occupation. 2003, 2013, and 2023 show the relative change from the prior",
    "period shown rather than the share itself. Net Change is the relative",
    "change in share from 1993 to 2023."
  )
)

save_table(tab_014, "tab_3_01", profile = "wide", stub_width_px = 260)
```

### Table 3.02 - How Concentrated Are Temporary Visa Holders in STEM Occupations 
```{r}
meta <- get_meta("TAB302")

short_names <- c(
  "Biological, agricultural, and other life scientists" = "Biological scientists",
  "Computer and mathematical scientists"                = "Computer/math scientists",
  "Physical and related scientists"                     = "Physical scientists",
  "Social and related scientists"                        = "Social scientists",
  "Engineers"                                            = "Engineers",
  "S&E-related occupations"                               = "S&E-related occupations",
  "Non-S&E occupations"                                   = "Non-S&E occupations"
)

row_order <- c("Computer/math scientists", "Engineers", "Biological scientists",
               "Physical scientists", "Social scientists",
               "S&E-related occupations", "Non-S&E occupations")

df_011 <- read_fig("TAB302") %>%
  mutate(across(-Year, as.numeric)) %>%
  pivot_longer(-Year, names_to = "Occupation", values_to = "Rate") %>%
  filter(!is.na(Rate), Year %in% c(2017, 2019, 2021, 2023)) %>%
  mutate(Rate = Rate * 100,
         Occupation = recode(Occupation, !!!short_names)) %>%
  pivot_wider(names_from = Year, values_from = Rate, names_prefix = "Y") %>%
  mutate(
    Chg2019 = Y2019 - Y2017,
    Chg2021 = Y2021 - Y2019,
    Chg2023 = Y2023 - Y2021,
    Change  = Y2023 - Y2017,
    Occupation = factor(Occupation, levels = row_order)
  ) %>%
  arrange(Occupation) %>%
  select(Occupation, Y2017, Chg2019, Chg2021, Chg2023, Y2023, Change)

TAB302 <- build_occupation_share_table(
  df_011, "TAB302", meta,
  level_cols = c("Y2017", "Y2023"),
  chg_cols   = c("Chg2019", "Chg2021", "Chg2023")
)

save_table(TAB302, "tab_3_02", profile = "wide", stub_width_px = 300)
```


### Table 3.03 - Top Employers Receiving H-1B Approvals
```{r tab_012}
meta <- get_meta("TAB303")

fig030_shares <- function() {
  raw <- read_fig("FIG302")
  names(raw) <- trimws(names(raw))

  total_row <- raw %>% filter(trimws(Company) == "Grand Total")
  totals <- total_row %>%
    select(-Company, -`Grand Total`) %>%
    pivot_longer(everything(), names_to = "Year", values_to = "Total") %>%
    mutate(Year = as.integer(Year), Total = as.numeric(Total))

  raw %>%
    filter(!trimws(Company) %in% c("Grand Total")) %>%
    mutate(Company = trimws(Company)) %>%
    select(-`Grand Total`) %>%
    pivot_longer(-Company, names_to = "Year", values_to = "Count") %>%
    mutate(Year = as.integer(Year),
           Count = as.numeric(replace_na(Count, 0))) %>%
    filter(Year != 2026) %>%                     # drop partial year
    left_join(totals, by = "Year") %>%
    mutate(Share = Count / Total * 100)
}

shares <- fig030_shares()

# proper-case helper (acronyms preserved, minor words lowercased)
proper_case <- function(x,
                        keep_upper = c("IBM", "HCL", "LLP", "LLC", "US", "USA", "EY"),
                        small = c("and", "of", "the", "for")) {
  vapply(x, function(s) {
    words <- str_split(s, "\\s+")[[1]]
    out <- vapply(seq_along(words), function(i) {
      w <- words[i]
      if (toupper(w) %in% keep_upper)             toupper(w)
      else if (w == "&")                          "&"
      else if (tolower(w) %in% small && i != 1)   tolower(w)
      else                                        str_to_title(w)
    }, character(1))
    paste(out, collapse = " ")
  }, character(1), USE.NAMES = FALSE)
}

# union of firms that were top-10 in EITHER year
union_firms <- shares %>%
  filter(Year %in% c(2017, 2025)) %>%
  group_by(Year) %>%
  slice_max(Count, n = 10) %>%
  ungroup() %>%
  distinct(Company) %>%
  pull(Company)

wide <- shares %>%
  filter(Year %in% c(2017, 2025), Company %in% union_firms) %>%
  select(Company, Year, Share) %>%
  pivot_wider(names_from = Year, values_from = Share, names_prefix = "Y")

df_012 <- wide %>%
  mutate(
    Company = proper_case(recode(Company, !!!short_names)),
    Y1 = Y2017, Y2 = Y2025,
    Change = if_else(Y2017 > 0, (Y2025 - Y2017) / Y2017 * 100, NA_real_)
  ) %>%
  arrange(desc(Change)) %>%          # order by 2025 share (today's leaders first)
  select(Company, Y1, Y2, Change)

tab_012 <- build_h1b_top10_table(df_012, "tab_012", meta,
                                 first_year = 2017, last_year = 2025)

save_table(tab_012, "tab_3_03", profile = "standard", stub_width_px = 360)
```


# Chapter 4

### Table 4.01 - Four in Ten STEM Entrants Finish with a STEM Degree — and That Barely Moved
```{r tab_018}
meta <- get_meta("TAB401")   # or a dedicated tab_038a entry

tab_038a <- build_bookend_table(
  name          = "TAB401",
  meta          = meta,
  first         = "2003/04",
  last          = "2011/12",
  id_col        = "cohort",
  spanner_label = "Share of STEM entrants",
  change_type   = "relative",
  recode_cats = c(
    stem_bachelors    = "Completed Bachelor's Degree in STEM Field",
    nonstem_bachelors = "Completed Bachelor's Degree in Another, Non-STEM Field",
    subbac_credential = "Completed A Subbac Credential",
    still_enrolled    = "Still Enrolled",
    left_no_degree    = "Left with No Degree"
  ),
  cat_order = c(
    "Completed Bachelor's Degree in STEM Field",
    "Completed Bachelor's Degree in Another, Non-STEM Field",
    "Completed A Subbac Credential",
    "Still Enrolled",
    "Left with No Degree"
  ),
  footnote = paste(
    "Values are the share of each entering STEM cohort ending in each outcome;",
    "each year column sums to 100%. Net change is the relative change from the",
    "2003/04 cohort to the 2011/12 cohort."
  )
)
save_table(tab_038a, "tab_4_01", profile = "standard", stub_width_px = 340)
```


### Table 4.02 - How Many Times Do STEM Majors Change Majors
```{r TAB402}
meta <- get_meta("TAB402")
TAB402 <- build_bookend_table(
  "TAB402", meta,
  first = "2004/2009 Survey", last = "2012/2017 Survey",
  id_col = "Survey",
  first_label = "2004/2009", last_label = "2012/2017",
  spanner_label = "Share of STEM entrants",
  change_type = "relative",
  recode_cats = c("Never"          = "Never switched",
                  "Once"           = "Switched once",
                  "More than Once" = "Switched 2+ times"),
  cat_order = c("Never switched", "Switched once", "Switched 2+ times"),
  footnote = paste(
    "Share of STEM entrants by number of major switches. Net change is the",
    "relative change from the 2004/2009 to the 2012/2017 cohort."
  )
)
save_table(TAB402, "tab_4_02")
TAB402
```

### Table 4.03 - What Do Americans Study Abroad?
```{r tab_016}
meta <- get_meta("TAB403")
tab_016 <- build_bookend_table(
  "TAB403", meta,
  first = 2005, last = 2024,
  spanner_label = "Share of students",
  change_type = "relative",
  sort_by = "change",
  recode_cats = c("STEM Fields"           = "STEM",
                  "Business & Management" = "Business & management",
                  "Social Sciences"       = "Social sciences",
                  "Fine and Applied Arts" = "Fine & applied arts"),
  flagged_rows = c("Social sciences", "Humanities", "Other"),
  flag_note = paste(
    "Open Doors revised its field taxonomy between 2013 and 2014; shares for",
    "these categories are not directly comparable across that break, so the",
    "reported change reflects reclassification as well as any real shift."
  ),
  footnote = paste(
    "Share of U.S. study-abroad students by field of study.",
    "Net change is the relative change from 2005 to 2024."
  )
)

save_table(tab_016, "tab_4_03", profile = "standard", stub_width_px = 260)
```

### Table 4.04 - Where Do American's Study Abroad
```{r tab_017}
meta <- get_meta("TAB404")
tab_017 <- build_bookend_table(
  "TAB404", meta,
  first = 2005, last = 2024,
  spanner_label = "Share of students",
  change_type = "relative",
  sort_by = "last",
  drop_cols = "Antarctica",
  recode_cats = c("Africa, Sub-Saharan"   = "Sub-Saharan Africa",
                  "Multiple Destinations" = "Multiple destinations"),
  footnote = paste(
    "Share of U.S. study-abroad students by host region. Net change is the",
    "relative change from 2005 to 2024, which is volatile for small-share",
    "regions (e.g., the Middle East & North Africa moved 1.0% to 1.6%)."
  )
)

save_table(tab_017, "tab_4_04", profile = "standard", stub_width_px = 280)
```


# Chapter 5 

### Table 5.01 - When Did China Catch At Each AI Conference
```{r table 019}
tab_501a <- build_conference_summary_table_split(
  name       = "tab_019a",
  meta       = get_meta("TAB501"),
  conf_names = conf_names,
  data_name  = "TAB501",
  row_slice  = 1:20
)

tab_501b <- build_conference_summary_table_split(
  name       = "tab_019b",
  meta       = get_meta("TAB501"),
  conf_names = conf_names,
  data_name  = "TAB501",
  row_slice  = 11:20
)

save_table(tab_501a, "tab_5_01a", profile = "ultrawide", stub_width_px = 155)
save_table(tab_501b, "tab_5_01b", profile = "ultrawide", stub_width_px = 155)
```

### Table 5.02 - How Many U.S. Universities Rank in the Global Top 100? 
```{r TAB502}
meta <- get_meta("TAB502")
num  <- function(x) parse_number(as.character(x))

field_order <- c("Physical Sciences", "Life Sciences", "Engineering", "Computer Science")
years_shown <- c(2014, 2018, 2022, 2026)

raw <- read_fig("TAB502") %>%
  rename(Year = 1) %>%
  mutate(Year = as.integer(num(Year)))

cs_2017 <- raw %>% filter(Year == 2017) %>% pull(`Computer Science`) %>% num()

df_001 <- raw %>%
  select(Year, all_of(field_order)) %>%                 # drop precomputed Average
  filter(Year %in% years_shown) %>%
  pivot_longer(-Year, names_to = "Field", values_to = "value") %>%
  mutate(value = num(value)) %>%
  pivot_wider(names_from = Year, values_from = value, names_prefix = "Y") %>%
  mutate(
    # computer science was first ranked in 2017; drop its 2017 count into the
    # 2014 slot so its change and net columns chain consistently
    Y2014   = if_else(Field == "Computer Science" & is.na(Y2014), cs_2017, Y2014),
    Chg2018 = Y2018 - Y2014,
    Chg2022 = Y2022 - Y2018,
    Chg2026 = Y2026 - Y2022,
    Change  = Y2026 - Y2014,
    Field   = factor(Field, levels = field_order)
  ) %>%
  arrange(Field) %>%
  select(Field, Y2014, Chg2018, Chg2022, Chg2026, Y2026, Change)

TAB502 <- build_ranking_delta_table(
  df_001, "TAB502", meta,
  level_cols  = c("Y2014", "Y2026"),
  chg_cols    = c("Chg2018", "Chg2022", "Chg2026"),
  change_type = "absolute",
  footnote = paste(
    "Counts are U.S. universities in the global top 100 in each field, in 2014 and",
    "2026. Interior columns show the change in that count from the prior shown year;",
    "Net Change is the change from 2014 to 2026."
  ),
  cell_footnote       = paste(
    "Computer science was first ranked in 2017; its 2017 count is shown here in the",
    "2014 column so its change and net columns can be computed."
  ),
  cell_footnote_field = "Computer Science",
  cell_footnote_col   = "Y2014"
)
save_table(TAB502, "tab_5_02")
TAB502
```

### Table 5.03 - Is R&D Spending Aimed at Today or Tomorrow? 
```{r table_020}
meta <- get_meta("TAB503")

type_labels <- c(
  "Basic Research"           = "Basic research (for tomorrow)",
  "Applied Research"          = "Applied research (for today)",
  "Experimental Development"  = "Experimental development (for today)"
)
type_order <- c("Basic research (for tomorrow)",
                "Applied research (for today)",
                "Experimental development (for today)")

df_125 <- read_fig("FIG509") %>%
  rename(Year = 1, Country = 2, Type = 3, value = 4) %>%
  mutate(Year  = as.numeric(as.character(Year)),
         value = as.numeric(value)) %>%
  filter(Year %in% c(1994, 2004, 2014, 2024)) %>%
  mutate(Type = recode(Type, !!!type_labels)) %>%
  pivot_wider(names_from = Year, values_from = value, names_prefix = "Y") %>%
  mutate(
    Chg2004 = (Y2004 - Y1994) / Y1994 * 100,
    Chg2014 = (Y2014 - Y2004) / Y2004 * 100,
    Chg2024 = (Y2024 - Y2014) / Y2014 * 100,
    Change  = (Y2024 - Y1994) / Y1994 * 100,
    Country = factor(Country, levels = c("United States", "China")),
    Type    = factor(Type, levels = type_order)
  ) %>%
  arrange(Country, Type) %>%
  select(Country, Type, Y1994, Chg2004, Chg2014, Chg2024, Y2024, Change)

tab_021 <- build_rd_type_delta_table(
  df_125, "tab_022", meta,
  level_cols = c("Y1994", "Y2024"),
  chg_cols   = c("Chg2004", "Chg2014", "Chg2024")
)

save_table(
  tab_021,
  "tab_5_03",
  profile = "ultrawide",
  stub_width_px = 300
)
```

### Table 5.04 - Business Now Dominates R&D in Both Countries 
```{r tab_022}
meta <- get_meta("TAB504")

sector_labels <- c(
  "Business enterprise sector" = "Business enterprise",
  "Government sector"          = "Government",
  "Higher education sector"    = "Higher education",
  "Private non-profit sector"  = "Private non-profit"
)
sector_order <- c("Business enterprise", "Government",
                  "Higher education", "Private non-profit")

df_022 <- read_fig("FIG510") %>%
  rename(Year = 1, Country = 2, Type = 3, value = 4) %>%
  mutate(Year  = as.numeric(as.character(Year)),
         value = as.numeric(value)) %>%
  filter(Type %in% names(sector_labels),
         Year %in% c(1994, 2004, 2014, 2024)) %>%
  mutate(Type = recode(Type, !!!sector_labels)) %>%
  complete(Country, Type, Year, fill = list(value = 0)) %>%   # China private non-profit -> $0
  pivot_wider(names_from = Year, values_from = value, names_prefix = "Y") %>%
  mutate(
    Chg2004 = (Y2004 - Y1994) / Y1994 * 100,
    Chg2014 = (Y2014 - Y2004) / Y2004 * 100,
    Chg2024 = (Y2024 - Y2014) / Y2014 * 100,
    Change  = (Y2024 - Y1994) / Y1994 * 100,
    across(c(Chg2004, Chg2014, Chg2024, Change), ~ ifelse(is.finite(.), ., 0)),  # 0/0 -> 0%
    Country = factor(Country, levels = c("United States", "China")),
    Type    = factor(Type, levels = sector_order)
  ) %>%
  arrange(Country, Type) %>%
  select(Country, Type, Y1994, Chg2004, Chg2014, Chg2024, Y2024, Change)

tab_022 <- build_rd_type_delta_table(
  df_022, "tab_022", meta,
  level_cols = c("Y1994", "Y2024"),
  chg_cols   = c("Chg2004", "Chg2014", "Chg2024"),
  footnote = paste(
    "The first and last columns are spending levels in billions of US dollars (PPP",
    "converted). The interior columns show the relative change from the prior shown",
    "year rather than the level. Net Change is the relative change across the full",
    "window. Sectors denote where R&D is performed. China reports no private non-profit",
    "R&D; it is shown as $0, and its relative-change cells (change from a zero base is",
    "undefined) appear as 0%."
  )
)

save_table(
  tab_022,
  "tab_5_04",
  profile = "ultrawide",
  stub_width_px = 300
)
```

### Table 5.05 - Who Patents in Critical Technologies
```{r table 020}
build_patents_table(
  name    = "tab_020",     # output id; adjust to next free table number
  meta    = get_meta("TAB505"),
  sources = c(
    "Biotechnology"       = "TAB505a",
    "Semiconductors"      = "TAB505b",
    "Computer technology" = "TAB505c"
  )
) %>%
  save_table("tab_5_05")
```

### Table 5.06 - Patent Leadership Shifted From U.S. and Japansse Incumbents to Asian Challengers
```{r TAB505}
meta <- get_meta("TAB506")
num  <- function(x) parse_number(as.character(x))

flag_slug <- c(
  "United States" = "united-states",
  "Japan"         = "japan",
  "Taiwan"        = "taiwan",
  "South Korea"   = "south-korea",
  "Germany"       = "germany",
  "China"         = "china"
)

df_023 <- read_fig("TAB506") %>%
  rename(Company = 1, Country = 2, Y2005 = 3, Y2015 = 4, Y2025 = 5) %>%
  mutate(across(c(Y2005, Y2015, Y2025), num)) %>%
  mutate(
    flag    = paste0("flags/", recode(Country, !!!flag_slug), ".png"),
    Chg2015 = Y2015 - Y2005,
    Chg2025 = Y2025 - Y2015,
    Change  = Y2025 - Y2005
  ) %>%
  arrange(desc(Change)) %>%
  select(Company, flag, Y2005, Chg2015, Y2015, Chg2025, Y2025, Change)

TAB505 <- build_patent_company_delta_table(
  df_023, "TAB505", meta,
  level_cols  = c("Y2005", "Y2015", "Y2025"),
  chg_cols    = c("Chg2015", "Chg2025"),
  change_type = "absolute"
)
save_table(TAB505, "tab_5_06", profile = "ultrawide", stub_width_px = 360)
```





# Chapter 6

### Table 6.01 - Do Stay Intentions Differ by Region of Origin? 
```{r table 028}
meta <- get_meta("TAB601")

tab_028 <- build_bookend_table(
  "TAB601", meta,
  first = 2002, last = 2024,
  spanner_label = "Intend to stay",
  scale = 100,
  drop_cols = "Region and country or economy unknown",
  change_type = "ppt"
)

save_table(tab_028, "tab_6_01")
tab_028
```


### Table 6.02 - What New STEM Phds Plan to Do After Graduations
```{r table 025}
meta <- get_meta("TAB602")

tab_025 <- build_bookend_table(
  "TAB602", meta,
  first = 2021, last = 2024,
  spanner_label = "Share with U.S. plans",
  sort_by = "change",
  scale = 100
)

save_table(tab_025, "tab_6_02")
```

### Table 6.03 - Where U.S. Awarded PhDs Work After Graduation
```{r table 027}
meta <- get_meta("TAB603")

TAB603 <- build_sector_bookend_table(meta)

save_table(TAB603, "tab_6_03")
TAB603
```

### Table 6.04 - How Many PERM Certifications Go Unusued
```{r table 029}
meta <- get_meta("TAB604")

TAB604 <- build_perm_expiry_table(meta)

save_table(TAB604, "tab_6_04")
```
### Table 6.05 - Employment-Based Green Card Wait Times by Nationality
```{r table 026}
meta <- get_meta("TAB605")

TAB605 <- build_eb_combined_table(meta)

save_table(TAB605, "tab_6_05")
TAB605
```


